Agda Meets Accelerate

### 2012-08-01 paper/main.tex
conclusion

>>>>

We have build an experimental Agda frontend for the Accelerate
language. The goal of this experiment was to explore potential uses of
dependently-typed programming for data-parallel languages.

At the moment, the outcome of the experiment is mixed. It is
successful, because we have been able to construct Agda functions for
a representative sample of Accelerate's functionality.
However, there was less scope for encoding extra information in the
dependent types than we had hoped for. Exact matching of array bounds
works, but results in restrictions (like the problems with
\texttt{zipWith} and filtering) that were not anticipated.

Exploiting algebraic properties did not work out in the intended way, mainly
because it boils down to asserting that some AST denotes an
associative function. However, these assertions cannot be proven: the
proof would have to apply the semantics to the AST, but the AST is an
abstract type in our implementation. An AST representation in Agda
might give us a better handle at this problem.

In some places, the Agda frontend is less dynamic than Accelerate. In
a number of places, Accelerate accepts a run-time value of type
\texttt{Exp sh} for a shape argument, where the Agda frontend requires
a value of type \texttt{Shape}. To address this problem, we would have
to include a \texttt{Shape}-indexed encoding of the \texttt{Shape}
type in the \texttt{Element} type so that we can describe the type of
an expression whose value has a certain shape.

Finally, the type translation of Agda's FFI has many shortcomings that
caused a number of problems for transporting information between Agda
and Haskell. One part of the problem is, unfortunately, the rich type
structure of Accelerate which already encodes much useful
information. An alternative, untyped (or less-typed) interface to
Accelerate would make the adaptation to an Agda frontend much simpler.
